---
  title: "Dynamical Critical Transitions"
author: "Leonardo A. Saravia"
date: "9/3/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Estimation of the crititical transitions of an species (birds) inside a dynamical habitat (forest)

In this model the birds reproduce inside the forest and to give birth search for an empty place in the neighborhood

```{r dynCritSearchForEmptyPatch}
require(readr)
require(dplyr)
require(ggplot2)

source("R/functions.r")

# wordl = 101
#
mdl <- read_netlogo_simul("Simulations/ContactProcessForestAsTurtles CriticalValueBirds1.5-3forest1.5-3_world101-table.csv")

# mdl <- bind_rows(mdl,read_netlogo_simul("ContactProcessForestAsTurtles CriticalValueBirds2.5-4forest1.5-1.85_world101-table.csv"))

#
# Plot the proportion of birds as a function of forest and birds birth rates
#
mdl1 <- mdl %>% group_by(birth_rate_birds,birth_rate_forest) %>% 
  summarize(sd_hp=sd(habitat_proportion),habitat_proportion=mean(habitat_proportion), birds_proportion=mean(birds_proportion) )

require(plotly)
plot_ly(data=mdl1, x=~birth_rate_birds, y=~birth_rate_forest, z=~birds_proportion, type="scatter3d", mode="markers", color=~birds_proportion, marker = list(size = 1))

#
# Estimate critical birds critical point 
#
birds_crit_size <- data_frame()
birds_crit_size <- bind_rows(birds_crit_size, estim_crit_point_size(mdl,101) %>% mutate(type="BirthSelection"))


#
# Plot the critical point estimated as a binomial model with the proportion of birds relative to the habitat 
# 
# Size of the model 251*251
#
# hab_size <- 101*101 
# mdl1 <- mdl %>% filter(birth_rate_birds %in% c(1.8,2.0,2.6,2.9)) %>% mutate(nhabitat= habitat_proportion*hab_size, nbirds=birds_proportion*hab_size, bird_hab_prop=nbirds/nhabitat) %>% mutate(bird_hab_prop=ifelse(bird_hab_prop>1,1,bird_hab_prop))
# 
# ggplot(mdl1 , aes(birth_rate_forest,bird_hab_prop,  colour=birth_rate_birds)) + geom_point(alpha=0.2) + theme_bw() +  scale_color_viridis_c(guide=FALSE) +  binomial_smooth() + facet_wrap(~birth_rate_birds)
# 

# ggsave("Figures/birds_proportion_critical.png",width=6)




#
#
# wordl = 201
#

mdl <- read_netlogo_simul("Simulations/ContactProcessForestAsTurtles CriticalValueBirds1.5-3forest1.5-3_world201-table.csv")

#
# Plot the proportion of birds as a function of forest and birds birth rates
#
mdl1 <- mdl %>% group_by(birth_rate_birds,birth_rate_forest) %>% 
  summarize(sd_hp=sd(habitat_proportion),habitat_proportion=mean(habitat_proportion), birds_proportion=mean(birds_proportion) )

require(plotly)
plot_ly(data=mdl1, x=~birth_rate_birds, y=~birth_rate_forest, z=~birds_proportion, type="scatter3d", mode="markers", color=~birds_proportion, marker = list(size = 1))

#
# Estimate critical birds critical point 
#
birds_crit_size <- bind_rows(birds_crit_size, estim_crit_point_size(mdl,201) %>% mutate(type="BirthSelection"))


#
# wordl = 51
#

mdl <- read_netlogo_simul("Simulations/ContactProcessForestAsTurtles CriticalValueBirds1.5-3forest1.5-3_world51-table.csv")

#
# Plot the proportion of birds as a function of forest and birds birth rates
#
mdl1 <- mdl %>% group_by(birth_rate_birds,birth_rate_forest) %>% 
  summarize(sd_hp=sd(habitat_proportion),habitat_proportion=mean(habitat_proportion), birds_proportion=mean(birds_proportion) )

require(plotly)
plot_ly(data=mdl1, x=~birth_rate_birds, y=~birth_rate_forest, z=~birds_proportion, type="scatter3d", mode="markers", color=~birds_proportion, marker = list(size = 1))

#
# Estimate critical birds critical point 
#
birds_crit_size <- bind_rows(birds_crit_size, estim_crit_point_size(mdl,51) %>% mutate(type="BirthSelection"))


#
# finite size scaling to determine pcrit for infinite lattices 
#

birds_crit_size <- bind_rows(birds_crit_size,
                             birds_crit_size %>% group_by(birth_rate_forest) %>% mutate(iSize=1/(size*size)) %>% 
                               do({
                                 model=lm(birth_rate_birds_c ~ iSize, data=.)
                                 data.frame(birth_rate_birds_c=predict(model,newdata=data.frame(iSize=0)),
                                            pcSE=predict(model,newdata=data.frame(iSize=0),se.fit=T)$se.fit)
                               }) %>% mutate(size=0))


ggplot(birds_crit_size %>% filter(size==0), aes(birth_rate_forest,birth_rate_birds_c,  colour=factor(size))) +  scale_color_viridis_d() + theme_bw() + geom_point(size=0.5) + #+ xlim(1.5,3) 
  geom_smooth( se = FALSE,size=0.3) 

ggplot(birds_crit_size, aes(birth_rate_forest,birth_rate_birds_c,  colour=factor(size))) +  scale_color_viridis_d() + theme_bw() + geom_point(size=0.5) #+ xlim(1.5,3) 
#  geom_smooth( se = FALSE,size=0.3) #+ facet_wrap(~size)


#
# Estimate for a wider range 
#
mdl <- read_netlogo_simul("Simulations/ContactProcessForestAsTurtlesCriticalValueBirds1-2forest3-4.5_w101_birthSel-table.csv")


birds_crit_size <- bind_rows(birds_crit_size, estim_crit_point_size(mdl,101) %>% mutate(type="BirthSelection") )

saveRDS(birds_crit_size, "birds_crit_size.rds")

birds_crit_size <- readRDS("birds_crit_size.rds")


rm(mdl,mdl1)
save.image()

```

The results showed that there is little difference between the critical points for different sizes, and the critical point for infinite lattices.

# Model with noSelection

In this model the birds reproduce inside the forest and to give birth they select a place at random in the neighborhood if the place is a forest site with no birds it success else it dies.

```{r dynCritNoSearchForEmptyPatch}
require(readr)
require(dplyr)
require(ggplot2)

source("R/functions.r")

# wordl = 101
#
mdl <- read_netlogo_simul("Simulations/ContactProcessForestAsTurtlesSelection CriticalValueBirds2.5-4.5forest2.5-4.5_w51_noSel-table.csv")

#
# Plot the proportion of birds as a function of forest and birds birth rates
#
mdl1 <- mdl %>% group_by(birth_rate_birds,birth_rate_forest) %>% 
  summarize(sd_hp=sd(habitat_proportion),habitat_proportion=mean(habitat_proportion), birds_proportion=mean(birds_proportion) )

require(plotly)
plot_ly(data=mdl1, x=~birth_rate_birds, y=~birth_rate_forest, z=~birds_proportion, type="scatter3d", mode="markers", color=~birds_proportion, marker = list(size = 1))

#
# Estimate critical birds critical point 
#

birds_crit_size <- bind_rows(birds_crit_size, estim_crit_point_size(mdl,51) %>% mutate(type="NoSelection") )

#
# World size 101
#
mdl <- read_netlogo_simul("Simulations/ContactProcessForestAsTurtlesSelection CriticalValueBirds2.5-4.5forest2.5-4.5_w101_noSel-table.csv")
birds_crit_size <- bind_rows(birds_crit_size, estim_crit_point_size(mdl,101) %>% mutate(type="NoSelection") )

mdl <- read_netlogo_simul("Simulations/ContactProcessForestAsTurtlesSelection CriticalValueBirds2.5-4.5forest2.5-4.5_w201_noSel-table.csv")
birds_crit_size <- bind_rows(birds_crit_size, estim_crit_point_size(mdl,201) %>% mutate(type="NoSelection") )

ggplot(birds_crit_size %>% filter(type=="NoSelection"), aes(birth_rate_forest,birth_rate_birds_c,  colour=factor(size))) +  scale_color_viridis_d() + theme_bw() + geom_point(size=0.5)

ggplot(birds_crit_size %>% filter(size==51), aes(birth_rate_forest,birth_rate_birds_c,  colour=factor(type))) +  scale_color_viridis_d(guide=FALSE) + theme_bw() + geom_point(size=0.5) + #+ xlim(1.5,3) 
  geom_smooth( se = FALSE,size=0.3) #+ facet_wrap(~size)

rm(mdl,mdl1)
save.image()
```

# Model with adult selection

In this version of the model if the forest patch die the adults move to an empty forest patch and also the parents select an empty forest patch

```{r dynCritNoSearchForEmptyPatch}
require(readr)
require(dplyr)
require(ggplot2)

source("R/functions.r")

# world = 101
#
mdl <- read_netlogo_simul("Simulations/ContactProcessForestAsTurtlesSelection CriticalValueBirds1.2-3forest1.2-3_w101_adSel-table.csv")


#
# Plot the proportion of birds as a function of forest and birds birth rates
#
mdl1 <- mdl %>% group_by(birth_rate_birds,birth_rate_forest) %>% 
  summarize(sd_hp=sd(habitat_proportion),habitat_proportion=mean(habitat_proportion), birds_proportion=mean(birds_proportion) )

require(plotly)
plot_ly(data=mdl1, x=~birth_rate_birds, y=~birth_rate_forest, z=~birds_proportion, type="scatter3d", mode="markers", color=~birds_proportion, marker = list(size = 1))

#
# Estimate critical birds critical point 
#

birds_crit_size <- bind_rows(birds_crit_size, estim_crit_point_size(mdl,101) %>% mutate(type="AdultSelection") )

```

# Run the model from R using nlrx 

- Previously I ran the model using behaviorSpace

```{r dynCritAdultSelectionnlrx}


require(nlrx)
require(readr)
require(dplyr)
require(ggplot2)

source("R/functions.r")

# Unix default NetLogo installation path (adjust to your needs!):
netlogopath <- file.path("/home/leonardo/NetLogo")
simfolder <- "/home/leonardo/Dropbox/Projects/DynamicForestExtinction"
modelpath <- file.path(simfolder, "ContactProcessForestAsTurtlesSelection.nlogo")
outpath <- file.path(simfolder,"Simulations")

# If not defined set the JAVA version of your local 
if(Sys.getenv("JAVA_HOME")==""){
  Sys.setenv(JAVA_HOME = "/usr/lib/jvm/java-11-openjdk-amd64")
  ## "/usr/lib/jvm/java-8-oracle"
}

nl <- nl(nlversion = "6.1",
         nlpath = netlogopath,
         modelpath = modelpath,
         jvmmem = 2048)

nl@experiment <- experiment(expname="CP_birds1-1.6_forest3-4.5_adSel",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="false",
                            idsetup="setup-center",
                            idgo="go",
                            runtime=1000,
                            metrics=c("habitat-proportion", "birds-proportion"),
                            variables = list("birth-rate-forest" = list(min=3, max=4.5, step=0.01 , qfun="qunif"),
                                             "birth-rate-birds" = list(min=1, max=1.6, step=0.01, qfun="qunif")), 
                            constants = list("world-width" = 101,
                                             "world-height" = 101,
                                             "death-rate-birds" = 1,
                                             "death-rate-forest" = 1,
                                             "birds-behavior" = "\"AdultSelection\""))
#
# Set a full factorial design 
#
nl@simdesign <- simdesign_ff(nl=nl,nseeds=5)

#nl@simdesign <- simdesign_lhs(nl=nl,nseeds=1,samples=10000,precision=3)

# run in Paralell 
#
require(future)
plan(multisession,workers=24)
require(tictoc)
tic()
results <- run_nl_all(nl,split = 20)
toc()
plan(sequential)

#
# Using BehaviorSpace 80.5 sims/min

# Write the output
#
setsim(nl, "simoutput") <- results 
write_simoutput(nl)

# Read the output to calculate critical points
#
mdl <- read_netlogo_simul("Simulations/CP_birds1-1.6_forest3-4.5_adSel_ff.csv",skip = 0)
birds_crit_size <- bind_rows(birds_crit_size, estim_crit_point_size(mdl,101) %>% mutate(type="AdultSelection") )

rm(mdl,mdl1,nl)
save.image()

ggplot(birds_crit_size %>% filter(size==101), aes(birth_rate_forest,birth_rate_birds_c,  colour=factor(type))) +  scale_color_viridis_d() + theme_bw() + geom_point(size=0.5)  #+ xlim(1.5,3) 
  # geom_smooth(method="loess", se = FALSE,size=0.3) #+ facet_wrap(~size)
ggsave("Figures/CriticalExtinctionNeigh1.png",width=6,height=5,units="in",dpi=600)

# Save data.frame with critical extinction data
#
saveRDS(birds_crit_size, "birds_crit_size.rds")

```

# Change dispersal


# Unit Testing 

- Check if the model with variable dispersal behaves equal than the fixed dispersal at dispersal-distance = 1

    - NoSelection checked
    - BirthSelection
    
  
```{r unitTestingModelnlrx}

#
# Model with fixed neighborhood = 4 closest cells 
#
modelpath <- file.path(simfolder, "ContactProcessForestAsTurtlesSelection.nlogo")

nl <- nl(nlversion = "6.1",
         nlpath = netlogopath,
         modelpath = modelpath,
         jvmmem = 2048)

nl@experiment <- experiment(expname="CP_UnitTest_birds1.9-5.0_forest2.5",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="true",
                            evalticks = 901:1000,
                            idsetup="setup-full",
                            idgo="go",
                            runtime=1000,
                            metrics=c("habitat-proportion", "birds-proportion"),
                            variables = list("birth-rate-forest" = list(values=c(2.5,2.5,2.5)),
                                             "birth-rate-birds" = list(values=c(1.9,2.1,5.0)),
                                             "birds-behavior" = list(values=c("\"AdultSelection\"", "\"BirthSelection\"", "\"NoSelection\""))
                                             ),
                            constants = list("world-width" = 101,
                                             "world-height" = 101,
                                             "death-rate-birds" = 1,
                                             "death-rate-forest" = 1)
                            )

nl@simdesign <- simdesign_distinct(nl=nl,nseeds = 1)

results <- run_nl_all(nl)
# Write the output
#
if( !file.exists("Simulations/CP_UnitTest_birds1.9-5.0_forest2.5_distinct.csv")){
  setsim(nl, "simoutput") <- results 
  write_simoutput(nl)
}

mdl <- read_netlogo_simul("Simulations/CP_UnitTest_birds1.9-5.0_forest2.5_distinct.csv",skip = 0)
mdl %>% group_by(birds_behavior) %>% summarize(habitat_proportion=median(habitat_proportion),birds_proportion=median(birds_proportion))

#  birds_behavior habitat_proportion birds_proportion
#   <chr>                       <dbl>            <dbl>
# 1 AdultSelection              0.549            0.304
# 2 BirthSelection              0.547            0.321
# 3 NoSelection                 0.549            0.234

test_mdl <- mdl %>% mutate(dispersal="fixed")

#
# Model with variable neighborhood = 4 closest cells 
#
modelpath <- file.path(simfolder, "ContactProcessForestAsTurtlesSelectionRadius.nlogo")

nl <- nl(nlversion = "6.1",
         nlpath = netlogopath,
         modelpath = modelpath,
         jvmmem = 2048)

nl@experiment <- experiment(expname="CP_RadiusTest_birds1.9-5.0_forest2.5",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="true",
                            evalticks = 901:1000,
                            idsetup="setup-full",
                            idgo="go",
                            runtime=1000,
                            metrics=c("habitat-proportion", "birds-proportion"),
                            variables = list("birth-rate-forest" = list(values=c(2.5,2.5,2.5)),
                                             "birth-rate-birds" = list(values=c(1.9,2.1,5.0)),
                                             "birds-behavior" = list(values=c("\"AdultSelection\"", "\"BirthSelection\"", "\"NoSelection\""))
                                             ),
                            constants = list("world-width" = 101,
                                             "world-height" = 101,
                                             "death-rate-birds" = 1,
                                             "death-rate-forest" = 1)
                            )

nl@simdesign <- simdesign_distinct(nl=nl,nseeds = 1)

results <- run_nl_all(nl)
setsim(nl, "simoutput") <- results 
write_simoutput(nl)
mdl <- read_netlogo_simul("Simulations/CP_RadiusTest_birds1.9-5.0_forest2.5_distinct.csv",skip = 0)
mdl %>% group_by(birds_behavior) %>% summarize(habitat_proportion=median(habitat_proportion),birds_proportion=median(birds_proportion))

test_mdl <- bind_rows(test_mdl, mdl %>% mutate(dispersal="variable"))

ggplot(test_mdl, aes(step,habitat_proportion,colour=dispersal)) +  scale_color_viridis_d() + theme_bw() + geom_point(size=0.5)  + facet_wrap(~birds_behavior)

ggplot(test_mdl, aes(step,birds_proportion,colour=dispersal)) +  scale_color_viridis_d() + theme_bw() + geom_point(size=0.5)  + facet_wrap(~birds_behavior)

test_mdl %>% group_by(birds_behavior,dispersal) %>% summarize(habitat_proportion=median(habitat_proportion),birds_proportion=median(birds_proportion))


#   birds_behavior dispersal habitat_proportion birds_proportion
#   <chr>          <chr>                  <dbl>            <dbl>
# 1 AdultSelection fixed                  0.549            0.304
# 2 AdultSelection variable               0.550            0.312
# 3 BirthSelection fixed                  0.547            0.321
# 4 BirthSelection variable               0.546            0.304
# 5 NoSelection    fixed                  0.549            0.234
# 6 NoSelection    variable               0.549            0.237

require(kSamples)

df <- test_mdl %>% group_by(birds_behavior) %>% do({
  ad <- ad.test(habitat_proportion ~ dispersal, data=.)
  data.frame(t(ad$ad[1,]))
  })
df <- test_mdl %>% group_by(birds_behavior) %>% do({
  ad <- ad.test(birds_proportion ~ dispersal, data=.)
  data.frame(t(ad$ad[1,]))

  })


rm(ad,df,mdl,nl,results)
save.image()
```