---
title: "Single large or Several Small SLOSS "
author: "Leonardo A. Saravia"
date: "9/3/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

needed_packages <- c(
    "tidyverse"
  , "lubridate"
  , "nlrx")

lapply(needed_packages, function(x) { if(!require(x,character.only = TRUE)) install.packages(x)} )

theme_set(theme_bw())
source("R/functions.r")


#### make changes
# Unix default NetLogo installation path (adjust to your needs!):
netlogopath <- file.path("/home/leonardo/NetLogo")
simfolder <- "/home/leonardo/Dropbox/Projects/ariadnaNL"
modelpath <- file.path(simfolder, "ariadnaNL.nlogo")
outpath <- file.path(simfolder,"Simulations")

# If not defined set the JAVA version of your local 
if(Sys.getenv("JAVA_HOME")==""){
  Sys.setenv(JAVA_HOME = "/usr/lib/jvm/java-11-openjdk-amd64")
  ## "/usr/lib/jvm/java-8-oracle"
}

nl <- nl(nlversion = "6.1.1",
         nlpath = netlogopath,
         modelpath = modelpath,
         jvmmem = 2048)


```

## Why do several small patches hold more species that few large patches?

## Simulaciones para ajuste usando Latin Hypercubic Sampling 

* Incursion by common, generalist species. 
* Minimun patch size selective extinction and nestedness
* Time since patch creation, accumulation of species losses
* Habitat heterogeneity environmental clumping patchy species distribution
* Dominance of colonization/inmigration
* Population variability spreading the risk

## Lambda = 4

* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )



```{r SLOSS_lambda4_simulations}
require(tidyverse)

source("R/functions.r")

# Neutral communities 
#
mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_NoSelection_pf06_hpf3-59_dd1-3.csv")
sloss_sim <- mdl

names(mdl)
#
# Plot for neutral communities 
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )


require(plotly)
plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()


mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

# Plot for hierarchical communities 
#
mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_Hierarchical_pf06_hpf3-59_dd1-3.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

#
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

# Plot for BirthSelection communities 
#
mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_BirthSelection_pf06_hpf3-59_dd1-3.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)


#
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

# check how many simulations 
mdl1 <-sloss_sim %>% count(habitat_patch_size,birds_dispersal_distance,birds_behavior)

```


## Lambda = 2

* For patch sizes 3 9 without connectivity between patches (dispersal 1.1 - 1.2 ) the patch populations get extinct 
* Here I added another field "Migration0_at_deforestation" 

```{r SLOSS_lambda4_simulations}
require(tidyverse)

source("R/functions.r")

# Neutral communities 
#
# No migration after deforestation
mdl <- read_netlogo_simul("Simulations/NoSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

## With migration
#
mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_NoSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

names(mdl)
#
# Plot for neutral communities 
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )


mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

# Plot for hierarchical communities 
#
mdl <- read_netlogo_simul("Simulations/Hierarchical_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_Hierarchical_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)


#
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()


# Plot for BirthSelection communities 
#

mdl <- read_netlogo_simul("Simulations/BirthSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

## with migration

mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_BirthSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

# Add Migration0_at_deforestation=FALSE to previous simulations 
#
sloss_sim <- sloss_sim %>% mutate( Migration0_at_deforestation = if_else(is.na(Migration0_at_deforestation), FALSE, Migration0_at_deforestation )) 


#
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>%
  mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

```



## Not fragmented communities Lambda = 1.7 - 4

```{r SLOSS_notfrag_simulations}
require(tidyverse)

source("R/functions.r")

# Not fragmented communities = habitat_patch_size = 100
#
mdl <- read_netlogo_simul("Simulations/NoHiBi_pf00_hpf3_dd1-3_lambda1.7-4.csv") %>% mutate(habitat_patch_size=100)

sloss_sim <- bind_rows(sloss_sim,mdl)

# Add Migration0_at_deforestation=FALSE to previous simulations 
#
#sloss_sim <- sloss_sim %>% mutate( Migration0_at_deforestation = if_else(is.na(Migration0_at_deforestation), FALSE, Migration0_at_deforestation )) #sloss_sim <- sloss_sim %>% mutate( habitat_patch_size = if_else(habitat_patch_size==0, 100, habitat_patch_size )) 

saveRDS(sloss_sim, "Simulations/sloss_sim.rds")

names(mdl)
#
# Plot for neutral communities 
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, birth_rate_birds,birds_behavior) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  ) 

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=birth_rate_birds)) + geom_point() + scale_color_viridis_d() + facet_wrap( ~birds_behavior )  + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=birth_rate_birds)) + geom_point() + scale_color_viridis_d() + facet_wrap( ~birds_behavior ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=count_birds,color=birth_rate_birds)) + geom_point() + scale_color_viridis_d() + facet_wrap( ~birds_behavior ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>%  mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=birth_rate_birds,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()+ facet_wrap( ~birds_behavior )

```

## Add more simulations lambda = 2 with Migration 

```{r SLOSS_combined_plots}

mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda2_Migrat.csv") 

sloss_sim <- bind_rows(sloss_sim,mdl)

save.image()

```


## Plots of combined behaviors and patch 100

```{r SLOSS_combined_plots}
require(tidyverse)

source("R/functions.r")

#mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation | (birth_rate_birds==2 & habitat_patch_size==100)) %>%  
mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

#mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3 || birds_dispersal_distance == 1.5) %>%  
mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")

ggsave("Figures/Species_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)


mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_shannon_diversity,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Shannon diversity") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")
ggsave("Figures/Shannon_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)
  

mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birth_rate_birds == 2 ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

#mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3 || birds_dispersal_distance == 1.5) %>%  
mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")

ggsave("Figures/Species_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)


mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_shannon_diversity,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Shannon diversity") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")
ggsave("Figures/Shannon_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)



```



```{r gen_fit_lhs, echo=FALSE, tidy=TRUE, message=FALSE, warning=FALSE, eval=FALSE}

nl@experiment <- experiment(expname="world201_Beta30_50_viaje0_3_trabajo4_8",
                            outpath=outpath,
                            repetition=1,
                            tickmetrics="true",
                            idsetup="setup",
                            idgo="go",
                            runtime=365,
                            metrics=c("count personas", "nro-hospitalizados", "nro-recuperados", "nro-fallecidos" ),
                              variables = list("Horas-en-viaje" = list(min=0, max=3, qfun="qunif"),
                                               "beta" = list(min=0.30, max=0.50, qfun="qunif"),
                                               "proporcion-hospitalizados"= list(min=0.20, max=0.45,qfun="qunif"),
                                               "Horas-en-trabajo"=list(min=4, max=8, qfun="qunif")
                                               ),                            
                              constants = list("world-width" = 201,
                                               "world-height" = 201,
                                               "infectados-iniciales" = 10,
                                               "max-personas-por-casa" = 10,
                                               "max-personas-por-trabajo" = 100,
                                               "periodo-latencia"=3.6,
                                               "Proporcion-asintomaticos"=0.43,
                                               "periodo-presintomatico"=1.5,
                                               "periodo-asintomatico"= 7.0,
                                               "periodo-hospitalizacion-fallecido" = 13.2,
                                               "periodo-hospitalizacion-recuperado" = 15,
                                               "Proporcion-fallecimiento-saturada"= 0.5,
                                               "Fallecido-sin-hospitalizacion"=0.18,
                                               "Proporcion-fallecimiento-hospitalizados"=0.09,
                                               "periodo-pre-hospitalizacion"= 2.5,
                                               "capacidad-de-camas"= 10000
                                             ))

#
# Set a latin hypercubic design 
#
nl@simdesign <- simdesign_lhs(nl=nl,
                               samples=300,
                               nseeds=4,
                               precision=2)

# run in Paralell 
#
require(future)
plan(multisession)
require(tictoc)
tic()
results <- run_nl_all(nl,split = 10)
toc()
plan(sequential)
names(results)

#
# Write the output
#
setsim(nl, "simoutput") <- results 
write_simoutput(nl)
```

